---
title: "R Notebook"
output: html_notebook
---

```{r message=FALSE}
library(tidyverse)
library(timetk)

wakefield_daily <- read_rds("00_data/processed/wakefield_daily_corrected_tbl.rds")
```

# Summary Statistics

```{r}
skimr::skim(wakefield_daily)
```

-   1888 days between 2019-03-01 and 2024-04-30

-   numeric response variable with no missing values

-   4 categorical variables (nominal)

# Exploratory Visualisation

## Distribution of appointments

### Total appointments

```{r}
# overall
total_appointments <- 
    wakefield_daily %>%
    summarise_by_time(
        .date_var = appointment_date,
        .by = "day",
        total_appointments = sum(count_of_appointments)
    )

total_appointments %>% 
    ggplot(aes(total_appointments)) +
    geom_histogram(bins = 20, fill = "skyblue", color = "grey30") +
    theme_bw() +
    labs(title = "Bimodal distribution due to zero values for holidays and weekends")

total_appointments %>% 
    filter(total_appointments > 0) %>% 
    ggplot(aes(total_appointments)) +
    geom_histogram(bins = 20, fill = "skyblue", color = "grey30") +
    theme_bw() +
    labs(title = "Approximately normal distribution when zero values omitted")

total_appointments %>% 
    filter(total_appointments > 0) %>%  
    select(total_appointments) %>% 
    summary()
```

### By Health Care Professional

#### GP

```{r}
gp_appointments <- 
    wakefield_daily %>%
    filter(hcp_type == "GP") %>% 
    summarise_by_time(
        .date_var = appointment_date,
        .by = "day",
        total_appointments = sum(count_of_appointments)
    )

gp_appointments %>% 
    filter(total_appointments > 0) %>% 
    ggplot(aes(total_appointments)) +
    geom_histogram(bins = 20, fill = "skyblue", color = "grey30") +
    theme_bw() +
    labs(title = "Approximately normal distribution when zero values omitted")

gp_appointments %>% 
    filter(total_appointments > 0) %>%  
    select(total_appointments) %>% 
    summary()
```

#### Other Practice Staff

```{r}
other_appointments <- 
    wakefield_daily %>%
    filter(hcp_type == "Other Practice staff") %>% 
    summarise_by_time(
        .date_var = appointment_date,
        .by = "day",
        total_appointments = sum(count_of_appointments)
    )

other_appointments %>% 
    filter(total_appointments > 0) %>% 
    ggplot(aes(total_appointments)) +
    geom_histogram(bins = 20, fill = "skyblue", color = "grey30") +
    theme_bw() +
    labs(title = "Approximately normal distribution when zero values omitted")

other_appointments %>% 
    filter(total_appointments > 0) %>%  
    select(total_appointments) %>% 
    summary()
```

## Total Appointments by Feature

```{r}
wakefield_daily %>%
    group_by(hcp_type) %>%
    summarise(total_appointments = sum(count_of_appointments)) %>%
    mutate(hcp_type = fct_reorder(hcp_type, -total_appointments)) %>%
    ggplot(aes(hcp_type, total_appointments)) +
    geom_col(fill = "skyblue", color = "grey30") +
    theme_bw()
```

```{r}
wakefield_daily %>%
    group_by(appt_mode) %>%
    summarise(total_appointments = sum(count_of_appointments)) %>%
    mutate(appt_mode = fct_reorder(appt_mode, -total_appointments)) %>%
    ggplot(aes(appt_mode, total_appointments)) +
    geom_col(fill = "skyblue", color = "grey30") +
    theme_bw()
```

```{r}
wakefield_daily %>%
    group_by(appt_status) %>%
    summarise(total_appointments = sum(count_of_appointments)) %>%
    mutate(appt_status = fct_reorder(appt_status, -total_appointments)) %>%
    ggplot(aes(appt_status, total_appointments)) +
    geom_col(fill = "skyblue", color = "grey30") +
    theme_bw()
```

```{r}
wakefield_daily %>%
    group_by(time_between_book_and_appt) %>%
    summarise(total_appointments = sum(count_of_appointments)) %>%
    ggplot(aes(time_between_book_and_appt, total_appointments)) +
    geom_col(fill = "skyblue", color = "grey30") +
    theme_bw() +
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5)) +
    xlab("Time between booking and appointment date")
```

### Findings

Most common levels are:

-   Other Practice Staff

-   Face to Face Appointments

-   Attended

-   Same Day

## Appointments vs. HCP Type by Feature

```{r}
wakefield_daily %>%
    filter(hcp_type == "GP" | hcp_type == "Other Practice staff",
           appt_mode %in% c("Face-to-Face", "Telephone")) %>% 
    group_by(hcp_type, appt_mode) %>%
    summarise(total_appointments = sum(count_of_appointments)) %>%
    mutate(appt_mode = fct_reorder(appt_mode, total_appointments)) %>%
    ggplot(aes(hcp_type, total_appointments, fill = appt_mode)) +
    geom_col(color = "grey30", position = "fill") +
    theme_bw() +
    tidyquant::scale_fill_tq()
```

```{r}
wakefield_daily %>%
    filter(hcp_type == "GP" | hcp_type == "Other Practice staff",
           appt_mode %in% c("Face-to-Face", "Telephone")) %>% 
    group_by(hcp_type, appt_status) %>%
    summarise(total_appointments = sum(count_of_appointments)) %>%
    mutate(appt_status = fct_reorder(appt_status, total_appointments)) %>%
    ggplot(aes(hcp_type, total_appointments, fill = appt_status)) +
    geom_col(color = "grey30", position = "fill") +
    theme_bw() +
    tidyquant::scale_fill_tq()
```

```{r}
wakefield_daily %>%
    filter(hcp_type == "GP" | hcp_type == "Other Practice staff") %>% 
    group_by(hcp_type, time_between_book_and_appt) %>%
    summarise(total_appointments = sum(count_of_appointments)) %>%
    mutate(prop_appointment = proportions(total_appointments)) %>% 
    ggplot(aes(hcp_type, prop_appointment, fill = time_between_book_and_appt)) +
    geom_col(color = "grey30", position = "dodge") +
    theme_bw() +
    tidyquant::scale_fill_tq()

wakefield_daily %>%
    filter(hcp_type == "GP" | hcp_type == "Other Practice staff") %>% 
    group_by(hcp_type, time_between_book_and_appt) %>%
    summarise(total_appointments = sum(count_of_appointments)) %>%
    mutate(hcp_type = fct_rev(hcp_type)) %>% 
    ggplot(aes(time_between_book_and_appt, total_appointments, fill = hcp_type)) +
    geom_col(position = "fill") +
    theme_bw() +
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5)) +
    xlab("Time between booking and appointment date") +
    tidyquant::scale_fill_tq()
```

```{r}
wakefield_daily %>%
    filter(hcp_type == "GP" | hcp_type == "Other Practice staff") %>%   
    mutate(appt_status = fct_relevel(appt_status, c("Attended", "Unknown", "DNA"))) %>%
    group_by(hcp_type, appt_status, time_between_book_and_appt) %>%
    summarise(total_appointments = sum(count_of_appointments)) %>%
    ggplot(aes(time_between_book_and_appt, total_appointments, fill = appt_status)) +
    geom_col(position = "fill") +
    theme_bw() +
    tidyquant::scale_fill_tq() +
    labs(title = "Very few DNA or Unknown status for same day appointments",
         subtitle = "DNA and Unknown proportion increases as delay increases") +
    facet_wrap(~ hcp_type) +
    theme(axis.text.x = element_text(angle = 90, hjust = 1))
    
```

### Findings

-   Higher proportion of overall appointments face to face for other staff compared to GPs

-   Higher proportion of DNA or unknown status for other staff compared to GPs

-   Higher proportion of same day appointments for GPs compared to other staff

-   Lower DNA rate for same day appointments - rate increases with time between booking and appointment

# Time Series Analysis

## Appointments attended

This provides an overview of the workload without including appointments not attended.

Attended = appointments delivered from surgery premises - ie. exclude home visits with status recorded as attended

```{r}
attended_daily_tbl <-
    wakefield_daily %>%
    filter(hcp_type == "GP" | hcp_type == "Other Practice staff", 
           appt_status == "Attended",
           appt_mode %in% c("Face-to-Face", "Telephone")) %>% 
    group_by(hcp_type) %>% 
    summarise_by_time(
        .date_var = appointment_date,
        .by = "day",
        appointments = sum(count_of_appointments)) %>% 
    pivot_wider(names_from = hcp_type, values_from = appointments) %>% 
    mutate(Total = GP + `Other Practice staff`) %>% 
    pivot_longer(cols = 2:4, names_to = "hcp_type", values_to = "appointments") %>% 
    mutate(hcp_type = fct_relevel(hcp_type, c("Total", "GP", "Other Practice staff")))
    

# by month
attended_daily_tbl %>% 
    group_by(hcp_type) %>% 
    summarise_by_time(
        .date_var = appointment_date,
        .by = "month",
        appointments = sum(appointments)) %>% 
    plot_time_series(.date_var = appointment_date,
                     .value = appointments,
                     .title = "Time Series Plot: Total Monthly Appointments")
# by week
attended_daily_tbl %>% 
    group_by(hcp_type) %>% 
    summarise_by_time(
        .date_var = appointment_date,
        .by = "week",
        appointments = sum(appointments)) %>% 
    plot_time_series(.date_var = appointment_date,
                     .value = appointments,
                     .title = "Time Series Plot: Total Weekly Appointments",
                     .smooth = FALSE)

# by day
attended_daily_tbl %>% 
    group_by(hcp_type) %>% 
    plot_time_series(.date_var = appointment_date,
                     .value = appointments,
                     .title = "Time Series Plot: Total Daily Appointments",
                     .smooth = FALSE)
```

Explore correlation between GP and Other Staff Appointments.

```{r}
attended_daily_tbl %>% 
    filter(hcp_type != "Total", appointments > 0) %>%
    pivot_wider(names_from = hcp_type, values_from = appointments) %>% 
    ggplot(aes(GP, `Other Practice staff`)) +
    geom_point(alpha = 0.5) +
    geom_smooth(method = 'lm') +
    theme_bw()

attended_daily_tbl %>% 
    filter(hcp_type != "Total", appointments > 0) %>%
    pivot_wider(names_from = hcp_type, values_from = appointments) %>% 
    select(2:3) %>% 
    cor()


```

Total GP appointments attended per day moderately strongly correlated to total Other Practice staff appointments (when zero values omitted)

GP workload is simpler to analyse because in this dataset "Other Practice Staff" is a homogeneous group that may contain HCA, PA, Nurse, Pharmacists etc. and may be better utilised as a feature rather than a response

GP appointments fell around pandemic then increased to a peak between 2022 and 2023. Values appear to be falling since 2023 . Total appointments increasing so reduced total gp workload occurring as workload for other practice staff increasing.

A further question to consider relates to workforce. What are appointment rates per GP? Can workforce data be used to estimate this? Is there any data that describes daily GP sessions?

## Appointments attended by Booking Type

Same day appointments appear to be the most frequent GP appointments and from my experience these are liable to the most fluctuation and impact on perceived workload and well-being.

```{r}
booking_type_attended_tbl <-
    wakefield_daily %>% 
    filter(hcp_type == "GP" | hcp_type == "Other Practice staff", 
           appt_status == "Attended",
           appt_mode %in% c("Face-to-Face", "Telephone", "Video Conference/Online"),
           time_between_book_and_appt != "Unknown") %>%
    mutate(booking_type = if_else(time_between_book_and_appt == "Same Day", "Same Day", "Advance")) %>%
    group_by(booking_type, hcp_type) %>%
    summarise_by_time(
        .date_var = appointment_date,
        .by = "day",
        appointments = sum(count_of_appointments)
    ) %>%
    ungroup()

glimpse(booking_type_attended_tbl)
```

Compare location of centre and variance between same day and advanced booked?

```{r}
total_attended_booking_type_tbl %>% 
    filter(appointments > 0) %>% 
    ggplot(aes(appointments)) +
    geom_density(aes(fill = booking_type), alpha = 0.7) +
    theme_bw()

total_attended_booking_type_tbl %>% 
    group_by(booking_type) %>% 
    filter(appointments > 0) %>% 
    summarise(
        mean = mean(appointments),
        var  = var(appointments),
        sd   = sd(appointments),
        ratio = mean/sd
    )

levene_test_result <- 
    car::leveneTest(appointments ~ booking_type, 
                    data = total_attended_booking_type_tbl %>% filter(appointments > 0))
levene_test_result
```

Statistically significant difference in variance between same day booked and advance booked appointments. This reflects my suspected theory around variability in same day demand.

```{r}
total_attended_booking_type_tbl %>% 
    group_by(booking_type) %>% 
    plot_time_series(.date_var = appointment_date, .value = appointments)
```

Overall same day booked appointments have remained fairly consistent with a slow approximately linear increase over the time series. Advance booked appointments drop dramatically in early 2020 which is consistent with the first UK lock down due to the covid pandemic.

## GP Appointments by Booking Type

```{r}
gp_attended_booking_type_tbl <-
    wakefield_daily %>% 
    filter(hcp_type == "GP", 
           appt_status == "Attended",
           appt_mode %in% c("Face-to-Face", "Telephone", "Video Conference/Online"),
           time_between_book_and_appt != "Unknown") %>%
    mutate(booking_type = if_else(time_between_book_and_appt == "Same Day", "Same Day", "Advance")) %>%
    group_by(booking_type) %>%
    summarise_by_time(
        .date_var = appointment_date,
        .by = "day",
        appointments = sum(count_of_appointments)
    ) %>%
    ungroup()
```

```{r}
gp_attended_booking_type_tbl %>% 
    filter(appointments > 0) %>% 
    ggplot(aes(appointments)) +
    geom_density(aes(fill = booking_type), alpha = 0.7) +
    theme_bw()

gp_attended_booking_type_tbl %>% 
    group_by(booking_type) %>% 
    filter(appointments > 0) %>% 
    summarise(
        mean = mean(appointments),
        var  = var(appointments),
        sd   = sd(appointments),
        ratio = mean/sd
    )

levene_test_result <- 
    car::leveneTest(appointments ~ booking_type, 
                    data = gp_attended_booking_type_tbl %>% filter(appointments > 0))
levene_test_result
```

```{r}
gp_attended_booking_type_tbl %>% 
    group_by(booking_type) %>% 
    plot_time_series(.date_var = appointment_date, .value = appointments)
```

## GP appointments vs. Appointment Mode by Booking Type

Homing in on the data further I want to explore how patterns of consulting are changing for GPs by comparing appointment mode with booking type.

```{r}
gp_appt_mode_vs_booking_tbl <-
    wakefield_daily %>% 
    filter(hcp_type == "GP", 
           appt_status == "Attended",
           appt_mode %in% c("Face-to-Face", "Telephone"),
           time_between_book_and_appt != "Unknown") %>%
    mutate(booking_type = if_else(time_between_book_and_appt == "Same Day", "Same Day", "Advance")) %>%
    group_by(booking_type, appt_mode) %>%
    summarise_by_time(
        .date_var = appointment_date,
        .by = "day",
        appointments = sum(count_of_appointments)
    ) %>%
    ungroup()
```

```{r}
gp_appt_mode_vs_booking_tbl %>% 
    filter(appointments > 0) %>%
    group_by(appt_mode, booking_type) %>% 
    summarise(
        mean = mean(appointments),
        var  = var(appointments),
        sd   = sd(appointments),
        ratio = sd/mean
    )
```

-   **Face-to-Face Appointments**: The same-day appointments have a higher absolute number (mean) even though their relative variability (CV) is lower. This indicates a higher workload overall, which can be taxing.

-   **Telephone Appointments**: The same-day appointments absolute number is not as high but there is higher relative and absolute variability, indicating volatility.

# Autocorrelation analysis

I am most interested in Same Day GP appointments broken down into total, face-to-face. and telephone.

## Total Same Day GP Appointments

```{r}
gp_appt_mode_vs_booking_tbl %>%
    summarise_by_time(
        .date_var = appointment_date,
        .by = "day",
        total_appointments = sum(appointments)
    ) %>% 
    plot_time_series(.date_var = appointment_date, .value = total_appointments)

gp_appt_mode_vs_booking_tbl %>%
    summarise_by_time(
        .date_var = appointment_date,
        .by = "day",
        total_appointments = sum(appointments)
    ) %>%
    plot_acf_diagnostics(.date_var = appointment_date, .value = total_appointments, .lags = 365)
    
```

## Same Day GP Appointments by Appointment Mode

```{r}
gp_appt_mode_vs_booking_tbl %>%
    group_by(appt_mode) %>% 
    summarise_by_time(
        .date_var = appointment_date,
        .by = "day",
        total_appointments = sum(appointments)
    ) %>% 
    plot_time_series(.date_var = appointment_date, .value = total_appointments)
```

```{r}
gp_appt_mode_vs_booking_tbl %>%
    group_by(appt_mode) %>% 
    summarise_by_time(
        .date_var = appointment_date,
        .by = "day",
        total_appointments = sum(appointments)
    ) %>%
    plot_acf_diagnostics(.date_var = appointment_date, .value = total_appointments, .lags = 365)
    
```

### Findings: 

# Exploring Seasonality

Focus on GP same day data - Totals vs Phone vs Face to face

```{r}

gp_appt_mode_vs_booking_tbl %>% 
    pivot_wider(names_from = appt_mode, values_from = appointments) %>% 
    mutate(Total = `Face-to-Face` + Telephone) %>% 
    pivot_longer(cols = 3:5, names_to = "appt_mode", values_to = "appointments")
```

```{r}
gp_appt_mode_vs_booking_tbl %>% 
    group_by(appt_mode) %>% 
    plot_seasonal_diagnostics(
        .date_var = appointment_date,
        .value = appointments,
        .feature_set = c("wday.lbl", "month.lbl", "year"),
        .title = ""
    )
```

```{r}



same_day_vs_booked_daily_tbl %>%
    filter(!appt_mode %in% c("Unknown", "Video Conference/Online")) %>%
    plot_seasonal_diagnostics(
        .date_var = appointment_date,
        .value = appointments,
        .feature_set = c("wday.lbl", "month.lbl", "year"),
        .title = "",
        .facet_vars = appt_mode,
        .interactive = FALSE
    ) +
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5))

wakefield_working_week_daily %>%
    filter(hcp_type == "GP" | hcp_type == "Other Practice staff",
           appt_mode %in% c("Face-to-Face", "Telephone")) %>% 
    mutate(day = wday(appointment_date, label = TRUE),
           appt_mode = fct_rev(appt_mode)) %>% 
    group_by(day, hcp_type, appt_mode) %>% 
    summarise(total_appointments = sum(count_of_appointments)) %>%
    ungroup() %>% 
    ggplot(aes(day, total_appointments, fill = appt_mode)) +
    geom_col(position = "fill") +
    theme_bw() +
    facet_wrap(~ hcp_type) +
    scale_fill_brewer(type = "qual", palette = 3)

wakefield_working_week_daily %>%
    filter(hcp_type == "GP" | hcp_type == "Other Practice staff",
           appt_mode %in% c("Face-to-Face", "Telephone")) %>% 
    mutate(month = month(appointment_date, label = TRUE),
           appt_mode = fct_rev(appt_mode)) %>% 
    group_by(month, hcp_type, appt_mode) %>% 
    summarise(total_appointments = sum(count_of_appointments)) %>%
    ungroup() %>% 
    ggplot(aes(month, total_appointments, fill = appt_mode)) +
    geom_col(position = "fill") +
    theme_bw() +
    facet_wrap(~ hcp_type) +
    scale_fill_brewer(type = "qual", palette = 3)

same_day_vs_booked_daily_tbl %>%
    filter(!appt_mode %in% c("Unknown", "Video Conference/Online")) %>%
    plot_seasonal_diagnostics(
        .date_var = appointment_date,
        .value = appointments,
        .feature_set = c("wday.lbl", "month.lbl", "year"),
        .title = "",
        .facet_vars = c(booking, hcp_type),
        .interactive = FALSE,
    ) +
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5))

same_day_vs_booked_daily_tbl %>% 
    mutate(day = wday(appointment_date, label = TRUE)) %>% 
    group_by(day, hcp_type, booking) %>% 
    summarise(appointments = sum(appointments)) %>%
    ungroup() %>% 
    ggplot(aes(day, appointments, fill = booking)) +
    geom_col(position = "fill") +
    theme_bw() +
    facet_wrap(~ hcp_type) +
    scale_fill_brewer(palette = 3)

same_day_vs_booked_daily_tbl %>% 
    mutate(month = month(appointment_date, label = TRUE),
           appt_mode = fct_rev(appt_mode)) %>% 
    group_by(month, hcp_type, booking) %>%
    summarise(appointments = sum(appointments)) %>%
    ungroup() %>% 
    ggplot(aes(month, appointments, fill = booking)) +
    geom_col(position = "fill") +
    theme_bw() +
    facet_wrap(~ hcp_type) +
    scale_fill_brewer(type = "qual", palette = 3)

```
